Tutorial: SQL-on-Hadoop Systems

### 2015-04-01 paper.tex
minor
adding drill

\documentclass{vldb}

%\usepackage{graphicx,amsmath,amssymb,balance}
%\usepackage{enumitem}
%\usepackage[ruled,vlined]{algorithm2e}
%\usepackage[usenames,dvipsnames]{color}
%\usepackage{ifthen}
%\usepackage{hyperref}	% package used for the \autoref* command
%\hypersetup{draft}	% don't create hyperlinks
%\usepackage{soul}
%\usepackage{mathtools}
%\usepackage{tikz}
%\usetikzlibrary{arrows}
\usepackage{balance}
\usepackage{url}

\begin{document}

\title{Tutorial: SQL-on-Hadoop Systems}

\numberofauthors{3}

% 1st author
\author{
\alignauthor
Daniel Abadi\\
\affaddr{Yale University}\\
\email{daniel.abadi@yale.edu}
\alignauthor
Fatma  {\"O}zcan\\
\affaddr{IBM}\\
\email{fozcan@us.ibm.com}
\alignauthor
Ippokratis Pandis\\
\affaddr{Cloudera}\\
\email{ippokratis@cloudera.com}
}

\maketitle

% ----------------
% --- Abstract ---
% ----------------
%\begin{abstract}
%\end{abstract}

\section{Introduction}

Over the last decade, the database field has witnessed significant major innovations and changes in enterprise data platforms. First came the wave of noSQL, which provide high scale-out systems, although at the expense of strong ACID transactions and declarative SQL processing. On the analytics side, Hadoop emerged as the platform for all analytics needs of the enterprise. Although Hadoop started with just MapReduce and HDFS, it evolved into a multi-framework system, which supports not only MapReduce but also Spark, Tez, SQL, and others. The driver that enables all these frameworks to work in a single cluster environment is the shared file system, HDFS.

Given the popularity of SQL in the enterprise, it was soon evident that SQL processing over HDFS data is critical in this new emerging enterprise data platform. In this tutorial, we use the term SQL-on-Hadoop to refer to systems that provide some level of declarative SQL(-like) processing over HDFS and noSQL data sources, using architectures that include computational or storage engines compatible with Apache Hadoop.

We classify SQL-on-Hadoop systems into six categories:

\begin{itemize}
\item{ {\bf Native systems with MPP architectures:} These are systems that employ an MPP database architecture that use stand-alone long running processing to execute SQL queries over HDFS data. This category includes systems such as impala, tajo, presto, and Big SQL}
\item {{\bf Native systems with big data run-times:} These are systems that utilize a big data run-time such as Tez, or Spark to execute SQL queries. This category includes systems such as Hive and Spark SQL.}
\item{{\bf Database hybrids:} These are systems that combine a relational SQL engine with Hadoop run-time. This category includes systems such as Hadapt, HAWQ, and Vortex.}
\item{ {\bf Database connectors:} This category includes the extensions provided by major database vendors to reach out to Hadoop and ingest HDFS data. } 
\item{ {\bf noSQL/Hbase specific variants:} These are systems that provide a declarative SQL layer on top of noSQL systems, HBase in particular. The systems in this category are Phoenix, and Splice Machine. Note that these systems focus on OLTP-like workloads, whereas all other categories address OLAP-like analytics}
\end{itemize}

An important aspect of SQL-on-Hadoop, which is orthogonal to all these categories, is the various HDFS file formats used to store the data. While text is an important format for in-situ SQL analysis, and ELT processing, various columnar formats have been proposed to provide fast analytics query processing. In this tutorial, we will first introduce these file formats, then we propose to describe the SQL-on-Hadoop systems in detail, examine their architectural differences, and cover in-depth their approach to indexing, updates, query optimization, as well as new algorithms they introduce in a 3-hour tutorial.

@@

\section{Category 1: Native systems with MPP architectures}

@@

\section{Category 2: Native systems with big data run-times}

Spark is a fast general purpose cluster computing engine that is compatible with Hadoop data.  There are three different systems that use Spark as their run-time for SQL processing: Shark \cite{sharksigmod13}, Hive on Spark \cite{hiveOnSpark}, and Spark SQL \cite{sparkSQL}. Shark replaces the MapReduce run-time of Hive \cite{hive} with Spark. It is still uses the Hive compiler, and interfaces but generates Spark jobs. Hive on Spark is another attempt at replacing MapReduce/Tez based run-times of Hive with Spark. Spark SQL uses the Shark \cite{sharksigmod13} run-time but replaces the HiveQL and compiler with SQL, and a new compiler that uses the catalyst query optimizer.

@@

\bibliographystyle{abbrv}
\bibliography{references}

\balance

\end{document}

>>>>

\documentclass{vldb}
%\usepackage{graphicx,amsmath,amssymb,balance}
%\usepackage{enumitem}
%\usepackage[ruled,vlined]{algorithm2e}
%\usepackage[usenames,dvipsnames]{color}
%\usepackage{ifthen}
%\usepackage{hyperref} % package used for the \autoref* command
%\hypersetup{draft} % don't create hyperlinks
%\usepackage{soul}
%\usepackage{mathtools}
%\usepackage{tikz}
%\usetikzlibrary{arrows}
\usepackage{balance}
\usepackage{url}
\begin{document}
\title{Tutorial: SQL-on-Hadoop Systems}

\numberofauthors{4}
% 1st author
\author{
\alignauthor
Daniel Abadi\\
\affaddr{Yale University}\\
\email{daniel.abadi@yale.edu}
\alignauthor 
\newcommand{\spmarcel}{\hspace{-8mm}}
\spmarcel Shivnath Babu\\
\spmarcel \affaddr{Duke University}\\
\spmarcel \affaddr{Durham, NC} \\
\email{\spmarcel shivnath@cs.duke.edu}
\alignauthor
Fatma {\"O}zcan\\
\affaddr{IBM Research}\\
\email{fozcan@us.ibm.com}
\alignauthor
Ippokratis Pandis\\
\affaddr{Cloudera}\\
\email{ippokratis@cloudera.com}
}
\maketitle
% ----------------
% --- Abstract ---
% ----------------
%\begin{abstract}
%\end{abstract}
\section{Introduction}
Enterprises are using Hadoop as a central data repository for all their data coming from various sources,  including operational systems, social media and the web, sensors and smart devices, as well as their applications. SQL processing over Hadoop data in particular gained significant traction over the couple of years, as many enterprise data management tools rely on SQL,
and many enterprise users are familiar and comfortable with it. As a result, the number of SQL-on-Hadoop systems have increased significantly.

%Over the last decade, the database field has witnessed significant major innovations and changes in enterprise data platforms. First came the wave of noSQL, which provide high scale-out systems, although at the expense of strong ACID transactions and declarative SQL processing. On the analytics side, Hadoop emerged as the platform for all analytics needs of the enterprise. Although Hadoop started with just MapReduce and HDFS, it evolved into a multi-framework system, which supports not only MapReduce but also Spark, Tez, SQL, and others. The driver that enables all these frameworks to work in a single cluster environment is the shared file system, Hadoop File System (HDFS).Given the popularity of SQL in the enterprise, it was soon evident that SQL processing over HDFS data is critical in this new emerging enterprise data platform.

In this tutorial, we use the term SQL-on-Hadoop to refer to systems that provide some level of declarative SQL(-like) processing over HDFS and noSQL data sources, using architectures that include computational or storage engines compatible with Apache Hadoop. We classify SQL-on-Hadoop systems into five categories:

\begin{itemize}
\item {{\bf Native systems with big data run-times:} These are systems that utilize a big data run-time such as Tez, or Spark to execute SQL queries. This category includes systems such as Hive and Spark SQL.}
\item{ {\bf Native systems with MPP architectures:} These are systems that employ an MPP database architecture that use stand-alone long running processing to execute SQL queries over HDFS data. This category includes systems such as impala, tajo, presto, and Big SQL}
\item{{\bf Database hybrids:} These are systems that combine a relational SQL engine with Hadoop run-time. This category includes systems such as Hadapt, HAWQ, and Vortex.}
\item{ {\bf Database connectors:} This category includes the extensions provided by major database vendors to reach out to Hadoop and ingest HDFS data. }
\item{ {\bf noSQL/Hbase specific variants:} These are systems that provide a declarative SQL layer on top of noSQL systems, HBase in particular. The systems in this category are Phoenix, and Splice Machine. Note that these systems focus on OLTP-like workloads, whereas all other categories address OLAP-like analytics}
\end{itemize}

An important aspect of SQL-on-Hadoop, which is orthogonal to all these categories, is the various HDFS file formats used to store the data. While text is an important format for in-situ SQL analysis, and ELT processing, various columnar formats have been proposed to provide fast analytics query processing. In this tutorial, we will first introduce these file formats, then we will discuss the characteristics, the pros and cons of each category in detail. In particular, we will examine their architectural differences, and cover in-depth their approach to indexing, updates, query optimization, as well as new algorithms they introduce in a 3-hour tutorial.

@@

\section{Category 1: Native systems with big data run-times}

Spark is a fast, general purpose cluster computing engine that is compatible with Hadoop data and addresses the shortcoming of MapReduce. There are three different systems that use Spark as their run-time for SQL processing: Shark \cite{sharksigmod13}, Hive on Spark \cite{hiveOnSpark}, and Spark SQL \cite{sparkSQL}. Shark replaces the MapReduce run-time of Hive \cite{hive} with Spark run-time. It is still uses the Hive compiler, and interfaces but generates Spark jobs. Shark introduced an in-memory columnar data organization that allowed for fast execution of SQL queries. Hive on Spark is another attempt at replacing MapReduce/Tez based run-times of Hive with Spark run-time, but still relies on the Hive interpreter. Finally, Spark SQL uses the Shark \cite{sharksigmod13} run-time but replaces the HiveQL and compiler with a new compiler for SQL and uses the catalyst query optimizer.

\section{Category 2: Native systems with MPP architectures}

As it became clear that the latency of launching jobs for each query is too expensive for interactive SQL query processing, there was a shift to shared-nothing database architectures for SQL processing over Hadoop data. In this group, a long-running process co-exists with dataNodes on each node in the cluster, and continuously answers SQL queries. The systems in this category implement their own run-times and data movements between their long-running processes.

After discussing the general characteristics of the systems in this category, we will examine 3 representatives systems in some detail: Impala, Big SQL, and Drill.

@@

Apache Drill is an open-source project which aims at providing SQL-like declarative processing over self-describing semi-structured data. Its focus is on analyzing data without emposing a fixed schema or creating tables in a catalog like Hive MetaStore. It runs queries over files and HBase tables, and discovers data when reading input data. For every fixed chunk of data, it discovers its schema, creates an in-memory columnar representation, and generates specific code for processing. As such, it can accommodate data chunks with varying schemas. Drill provides a columnar execution engine that delays materialization as late as possible.

@@

An important category of SQL-on-Hadoop includes systems that provide some level of SQL support over HBase data. HBase provides auto-sharding and fail over technology for scaling tables across multiple servers. It also enables updates, running on top of the HDFS, which itself does not support updates. HBase scales out to petabytes of data easily over a cluster of commodity hardware. It has become very popular to store IoT (Internet of Things) data, web, mobile and social data, as well as machine-generated data, as HBase supports high data ingestion rates that is typical for these data sets. In this category, we will examine two representative systems; namely Splice Machine \cite{splice} and Phoenix \cite{phoenix}.

Splice Machine \cite{splice} provides SQL support over HBase data using Apache Derby, targeting both operational as well as analytical workloads. It replaces the storage system and run-time of Derby with HBase and Hadoop. Splice Machine leverages the Derby compiler stack to generate execution plans that access HBase servers. Splice Machine adapts the Derby optimizer and planner to leverage HBase capabilities.

\bibliographystyle{abbrv}
\bibliography{references}
\balance
\end{document}
